---
title: "Reproduceable Reaearch Project 1"
author: "MayHsu"
date: "3 February 2019"
output: html_document
---

## 1. Code for reading in the dataset and/or processing the data
Read activity data set

```{r}
activity <- read.csv("activity.csv",header=TRUE,sep=",",dec=",")
```

## 2. Histogram of the total number of steps taken each day

Prepare the aggregated Total Steps per Date dataset by removing **NA** values and summerized.

```{r clean}
clean <- activity[!is.na(activity$steps),]
totalStepsPerDate <- aggregate(clean$steps ~ clean$date, FUN=sum )
colnames(totalStepsPerDate)<- c("Date", "Steps")
```

Show histogram of the total number of steps taken each day

```{r , echo=FALSE}
hist(totalStepsPerDate$Steps, xlab="Steps", main = "Total Steps per Day")
```

## 3. Mean and median number of steps taken each day


```{r}
meanSteps <- mean(totalStepsPerDate$Steps)
medianSteps <- median(totalStepsPerDate$Steps)
```

**Mean and Median ** number of steps taken each day are `r meanSteps` and `r medianSteps`.

## 4. Time series plot of the average number of steps taken

Prepare the aggregated Total Steps per Interval dataset.

```{r}
avgStepsPerInterval <- aggregate(clean$steps,by=list(clean$interval),mean)
colnames(avgStepsPerInterval) <- c("Interval","AvgSteps")
```

Show Time series plot of the average number of steps taken

```{r , echo=FALSE}
plot(avgStepsPerInterval,xlab = "Interval",ylab = "Average Steps",main = "Average Daily Activity Pattern",type="l")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
maxSteps <- max(avgStepsPerInterval$AvgSteps)
maxInterval <- avgStepsPerInterval[avgStepsPerInterval$AvgSteps == maxSteps,1]
```

The interval associated with maximum activity is interval `r maxInterval`, at `r maxSteps` steps.

## 6. Code to describe and show a strategy for imputing missing data

My strategy is to calculate average steps per interval first.
After that fill the missing data with calculated average steps by looping and updating one by one.

```{r}
StepsPerInterval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
for(i in 1:dim(activity)[1]){
        avgSteps <- StepsPerInterval[i]
        activitySteps <- activity[i,1]
        if(is.na(activitySteps)){
                activity[i,1] <- avgSteps
        }
}
```
## 7. Histogram of the total number of steps taken each day after missing values are imputed

Prepare the aggregated Total Steps per Date data set by using filled dataset.

```{r}
totalStepsPerDate_fill <- aggregate(activity$steps ~ activity$date, FUN=sum )
colnames(totalStepsPerDate_fill)<- c("Date", "Steps")
```

Show full picture histogram of the total number of steps taken each day after missing values are imputed

```{r , echo=FALSE}
hist(totalStepsPerDate_fill$Steps, xlab="Steps", main = "Total Steps per Day")
```


```{r}
revMeanSteps <- mean(totalStepsPerDate_fill$Steps)
revMedianSteps <- median(totalStepsPerDate_fill$Steps)
```

The original **Mean and Median ** number of steps taken each day are `r meanSteps` and `r medianSteps`.
Whereas the revised **Mean and Median ** number of steps taken each day are `r revMeanSteps` and `r revMedianSteps`.
So the mean value remains the same as prior to imputation, while the median value increased slightly.

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Prepare weekend and weekday data to use in plot.

```{r}
## Add one more day column to insert weekned and weekday data.
activity$day <- ifelse(weekdays(as.Date(activity$date)) == "Saturday" | weekdays(as.Date(activity$date)) == "Sunday", "weekend", "weekday")

# Calculate average steps per interval for weekends
avgStepsPerInterval_weekend <- aggregate(as.integer(activity[activity$day == "weekend",1]) ~ activity[activity$day == "weekend",3],FUN=mean)
colnames(avgStepsPerInterval_weekend) <- c("Interval","AvgSteps")

# Calculate average steps per interval for weekdays
avgStepsPerInterval_weekday <- aggregate(as.integer(activity[activity$day == "weekday",1]) ~ activity[activity$day == "weekday",3],FUN=mean)
colnames(avgStepsPerInterval_weekday) <- c("Interval","AvgSteps")
```

Show Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r , echo=FALSE}
par(mfrow=c(1,2))

plot(avgStepsPerInterval_weekend,xlab = "Interval",ylab = "Average Steps",main = "Average Activity Pattern(Weekend)",type="l")
plot(avgStepsPerInterval_weekday,xlab = "Interval",ylab = "Average Steps",main = "Average Activity Pattern(Weekday)",type="l")
```

---
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
---
